---
title: "lab_03"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## TASK: Load libraries

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse
library(tidyverse)

# Load janitor
library(janitor)
```

## TASK: Load data and clean up the column names

Load the two datasets we worked with in the pre_lab that are in the data folder: the Maryland payments to vendors and Maryland grants and loans AND, using janitor, change the column names so that they are lower-cased and any spaces are replaced with underscores.

```{r}
# Load in the Maryland state agency payments data, making sure that it correctly sets the datatypes using all of the rows
payments <- read_csv("data/State_of_Maryland_Payments_Data__FY2008_to_FY2024.csv", guess_max=369008)

# Load in the Maryland grants and loan data
md_grants_loans <- read_csv("data/State_of_Maryland_Grant_and_Loan_Data__FY2009_to_FY2022.csv")
```

## Answer questions

Q1: You want to know how much money state agencies have paid to the Prince George's County School system, but to do that you'll need to find all of the variations of how the school system might be listed in the payments data. Look at & search the records in the dataframe to start your search, starting with a broad search and then narrowing it down. Then write code to find rows where the vendor name contains some version of the county's school system.

HINT: you should use filter() just once, and inside your filter you want to look for the variations one at a time. Think "this version OR that version."

When you have found as many as possible, save the result to a new dataframe called `pg_schools`. How many records did you find? 

A1: 70 records

```{r}
cleaned_payments <- payments |>
  clean_names()

pg_schools <- cleaned_payments |> 
  filter(
    vendor_name == "PRINCE GEORGE'S COUNTY PUBLIC SCHOOLS" |
    vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOL" |
    vendor_name == "PRINCE GEORGES COUNTY PUBLIC SCHOOLS" |
    vendor_name == "PRINCE GEORGES COUNTY SCHOOL" |
    vendor_name == "PG COUNTY PUBLIC SCHOOL" | 
    vendor_name == "PRINCE GEORGES PUBLIC SCHOOLS"
  )
```

Q2: Using that `pg_schools` dataframe, write code to calculate the total amount of payments to the Prince George's County School system for each fiscal year. Which fiscal year had the largest total amount? Are there any aspects of the results that stand out to you?

A2:

```{r}
pg_schools |> 
  group_by(fiscal_year) |> 
  summarize(total = sum(amount)) 
```

Q3. You're interested in examining the vendors paid for capital projects at the University of Maryland. Search the payments data to find those payments, then write code to produce a list of each vendor and the total amount it received.

Look carefully at that list - what cleaning will it require to ensure that you have only one row per vendor and zip code? Provide at least three examples of vendor names that are similar and should be combined into one. Describe in words how you might clean that data so that any nearly identical vendors would look identical - what would you change and how?

A3. You would need to sort through the company names to see if there is any overlapping companies that have two separate rows. You can use case_when to find any duplicate names and group them together with one universal name. The three examples of similar vendor names are PLANO-COUDON LLC and PLANO-CONDON, LLC, WHITING TURNER CONTRACTING CO and WHITING TURNER CONTRACTING COMPANY, and THE BALLINGER CO and BALLINGER.

```{r}
umd_capital_projects <- cleaned_payments |> 
  filter(agency_name == "CAPITAL PROJECTS- UNIVERSITY OF MARYLAND")

umd_capital_projects |> 
  group_by(vendor_name) |> 
  summarize(total = sum(amount)) |> 
  arrange(desc(total))
```

Q4: You want to understand how much state grant money went to stem cell research projects each fiscal year. Search the data to find grants that fit that description. Once you find which column(s) contain the information you want, your code should find just the grants for stem cell research in a single filter. Then calculate the total amount of money for each fiscal year. Write a sentence summarizing how stem cell research funding has changed over time.

A4: The stem cell research funding has drastically increased over time, evident by 2020-2023 being the top four year totals. Back in 2010, the stem cell research funding only received 7870123 dollars but in 2023 the funding spiked to 12415121.

```{r}
cleaned_md_grants_loans <- md_grants_loans |>
  clean_names() |> 
  mutate(grantor = str_to_lower(grantor), grantee = str_to_lower(grantee), description = str_to_lower(description))

stem_cell_research <- cleaned_md_grants_loans |> 
  filter (
    description == "stem cell research grant" |
    description == "stem cell research grant, pre-commercialization research"
  )

stem_cell_research |> 
  group_by(fiscal_year) |> 
  summarize(total = sum(amount)) |> 
  arrange(desc(fiscal_year))
```

Q5: Using the Maryland grants and loan data, find records where the zip code is the one where the University of Maryland-College Park is located in and produce a list of grantees from that zip code, along with the total amount they received. Write a paragraph explaining how many grantees are there, what you would need to do to clean the results so that you had one record per grantee and which grantees are most interesting to you?

A5: There are 47 different grantees and a ton of them overlap with each other. Using case_when again would be best to structure all the grantee names to the exact same thing. It would turn all university of md college pa to the full name and make it much more organized to sift through. An interesting grantee to me is just someone's name (Elizabeth Gallauresi), which is weird because she works in Silver Spring and that is not the same zip code as College Park. There are just some random companies and restaurants in College Park that are receiving grants.

```{r}
college_park <- cleaned_md_grants_loans |> 
  filter (zip_code == 20740 | zip_code == 20742)

college_park |> 
  group_by(grantee) |> 
  summarize(total = sum(amount)) |> 
  arrange(desc(total))
```
