---
title: "lab_02"
author: "Derek Willis"
date: "2025-01-16"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load libraries and establish settings

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with the UMD courses and Maryland campaign expense data we used in the pre_lab.

```{r}
umd_courses <- read_rds("data/umd_courses.rds")
maryland_expenses <- read_csv("data/maryland_expenses.csv")
```

## Answer questions

**Q1.** How many departments have offered the courses with "Climate" in the title and at least one seat? Which department has offered the most? **A1.**

```{r}
umd_courses |> 
  filter(str_detect(title, "Climate") & seats >= 1) |> 
  group_by (department) |> 
  summarise(course_count = n()) |> 
  arrange(desc(course_count))
```

**Q2.** Of the departments listed in A1, which one is the most interesting to you, and why? **A2.**

The total number of University Honors classes with the keyword "Climate" is the most surprising to me. I didn't expect to see that as the third highest department and I definitely didn't expect it to have 21 different classes. I also thought some of the other departments would have more, like the Marine-Estuarine-Environmental Sciences and Agricultural and Resource Economics

**Q3** Which departments offered the most courses in the Fall 2024 term? And which departments in that term have at least one syllabus posted? Answer both of those questions and then compare the top five departments - how are they different?

**A3.** Business and Management had the most classes in the Fall of 2024. While the first three departments had a reasonable number of syllabus posted, the English department and Teaching and Learning, Policy and Leadership departments only had 11 and six syllabus posted for a total number of classes over 100.

```{r}
umd_courses |> 
  filter (term == 202408) |> 
  group_by(department) |> 
  summarise (
    course_count = n(),
    syllabus_count = sum(syllabus_count >=1)
  ) |> 
  arrange(desc(course_count))
```

**Q4.** Using the Maryland campaign expenses data, change the datatype of the expenditure_date column so that it is a date and not a chr, saving the result as a new dataframe. Then create a new column for the month based on expenditure_date. Then write code that shows total amount spent each month by Wes Moore's committee (you'll need to search the data to find the exact name, and we're not looking in payee_name). Write a sentence describing the pattern of spending over time. **A4.**

The total amount of money spent by Wes Moore's committee slowly increased over time, likely because he was accumulating more money from donations as the campaign went along.

```{r}
maryland_new_expenses <- maryland_expenses |> 
  mutate(
    expenditure_date= mdy(expenditure_date), 
    month = floor_date(expenditure_date, "month")
  ) |> 
  filter(committee_name == "Moore  Wes For Maryland") |> 
  group_by(month) |>
  summarise(total_amount = sum(amount))

```

**Q5.** Using case_when(), create a column in the Maryland expenses data called `spending_location` indicating whether each record spent money in Maryland or outside Maryland. For records that are in Maryland, make the new column's value "In-state" and for all others, make it "Out of state". Then write code that shows the total amount of money spent in each category and describe the results. You can do this in one statement or break it up into two statements (saving the new column to your dataframe). **A5**

The amount of money spent in-state was over the double the amount out of state.

```{r}
maryland_expenses <- maryland_expenses |>
  mutate(
    spending_location = case_when(
      str_detect(address, " Maryland ") ~ "In-state",
      str_detect(address, " California ") ~ "Out of state",
      str_detect(address, " Washington ") ~ "Out of state",
      str_detect(address, " Louisiana ") ~ "Out of state",
      str_detect(address, " Florida ") ~ "Out of state",
      str_detect(address, " North Carolina ") ~ "Out of state",
      str_detect(address, " Massachusetts ") ~ "Out of state",
      str_detect(address, " West Virginia ") ~ "Out of state",
      str_detect(address, " Virginia ") ~ "Out of state",
      .default = NA
    )
  )

maryland_expenses |>
  group_by(spending_location) |>
  summarize(total = sum(amount)) |>
  arrange(desc(total))

```

**Q6.** Choose your own adventure: using either the UMD courses or campaign expenses data, ask and answer a question that is different than the previous ones. You can use one of the ones you provided in the pre_lab. **A6.**

```{r}
maryland_expenses |> 
  group_by(committee_name) |>
  summarise(total_collected = sum(amount)) |>
  arrange(desc(total_collected))
```
Wes Moore collected the most amount of money by a significant margin, nearly doubling David Blair.
